---
title: "Why does Techstars think this ingredient is the most important one in a startup?"
author: "Ran Du"
date: "21/02/2020"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data used in this website is from Crunchbase datasets
## Step 1 we will read all the datasets about the startups.
```{r}
library(dplyr)
company1 <- read.csv("company_sea_2010.csv", header = T, stringsAsFactors = F)
company2 <- read.csv("company_sea_2011to2012.csv", header = T, stringsAsFactors = F)
company3 <- read.csv("company_sea_2013to2014.csv", header = T, stringsAsFactors = F)
company4 <- read.csv("company_sea_2015to2018.csv", header = T, stringsAsFactors = F)
company5 <- read.csv("company_sea_2019.csv", header = T, stringsAsFactors = F)

# Error message during bind_rows showed thats Numer.of.News.Articles that has comma in numeric number values
company1$Number.of.Articles <- as.numeric(gsub(",", "", company1$Number.of.Articles))
company2$Number.of.Articles <- as.numeric(gsub(",", "", company2$Number.of.Articles))
company3$Number.of.Articles <- as.numeric(gsub(",", "", company3$Number.of.Articles))
company4$Number.of.Articles <- as.numeric(gsub(",", "", company4$Number.of.Articles))
company5$Number.of.Articles <- as.numeric(gsub(",", "", company5$Number.of.Articles))

# Combine them together
companies <- bind_rows(company1, company2, company3, company4, company5)

# Looks like there are some capital firms or venture capital companies in the company that skewed the data
companies <- companies[(!grepl("Venture Capital",companies$Industries)),]
companies <- companies[(!grepl("Private Equity",companies$Industries)),]
companies <- companies[(!grepl("Capital",companies$Organization.Name)),]

# A lot of companies do not have founder info, therefore useless for this analysis. We would remove these companies.
companies <- companies[!is.na(companies$Number.of.Founders),]

# A lot of rounds of funding and funding amount are missing too. We will replace them with 0.
companies$Number.of.Funding.Rounds <- replace(companies$Number.of.Funding.Rounds, is.na(companies$Number.of.Funding.Rounds),0)
companies$Total.Funding.Amount<- replace(companies$Total.Funding.Amount, is.na(companies$Total.Funding.Amount),0)
companies$Total.Funding.Amount.Currency..in.USD.<- replace(companies$Total.Funding.Amount.Currency..in.USD., is.na(companies$Total.Funding.Amount.Currency..in.USD.),0)

# A lot of number of articles has missing values too
companies$Number.of.Articles <- replace(companies$Number.of.Articles, is.na(companies$Number.of.Articles),0)

# Created a founded year variable
library(lubridate)
companies$Founded.Year <- year(companies$Founded.Date)
```

## Let's start with some overview of funding 
```{r}
library(ggplot2)
library(scales)
ggplot(companies, aes(x = as.factor(Founded.Year), y = Total.Funding.Amount.Currency..in.USD.)) + 
    geom_jitter(col = "forestgreen", alpha = 0.4, size = 2.5) +
    scale_y_continuous(labels = dollar_format(scale = 1/1000000, suffix = "M")) + 
    xlab("Founded Year") +
    ylab("Total Funding to Date in USD") +
    ggtitle("Greater Seattle Area Startups 2010-2019") +
    geom_hline(yintercept = 5000000, color = "black", linetype = "dashed") +
    theme(plot.title = element_text(size = 10, face = "bold"), axis.title = element_text(size = 8), axis.text = element_text(size = 7.5))

ggsave("seattle_startups_funding_overview.jpg")
```

### Who are these top 10 companies?
```{r}
top_10_companies <- top_n(companies,10,Total.Funding.Amount.Currency..in.USD.)
top_10_companies <- top_10_companies[order(top_10_companies$Total.Funding.Amount.Currency..in.USD., decreasing = T),]
rownames(top_10_companies) <- 1:nrow(top_10_companies)
head(top_10_companies, 10)[,c(1,18,28)]
```

## Since top 10 companies will skew the data. Let's remove them for the rest of the analysis
## We will also create a aggregated table for median funding because median is a better metrics to look at when data is skewed
```{r}
companies <- companies %>% filter(!Organization.Name %in% top_10_companies$Organization.Name)

# Aggregate number of founders to 1,2,3, and 4+ and create a new variable for number of founder category
companies$Categroy.Number.of.Founders <- as.factor(ifelse(companies$Number.of.Founders == 1, "Solo", ifelse(companies$Number.of.Founders == 2, "Dynamic Duo", ifelse(companies$Number.of.Founders == 3, "Three Musketeers", "Fantastic Four and More"))))
companies$Categroy.Number.of.Founders <- factor(companies$Categroy.Number.of.Founders, c("Solo","Dynamic Duo", "Three Musketeers","Fantastic Four and More"))

# Create aggregated median table
median_funding1 <- aggregate(Total.Funding.Amount.Currency..in.USD.~ Founded.Year + Categroy.Number.of.Founders, data = companies, median)
```

## Step 2 - Data Visualization to examine the relationship between number of founders and funding
```{r}
ggplot(companies, aes(x= Number.of.Founders, y = log10(Total.Funding.Amount.Currency..in.USD.))) +
  geom_jitter(size = 0.7, color = "forestgreen") +
  geom_smooth(method = "lm", color = "black", lwd = 0.8) +
  facet_wrap(.~as.factor(Founded.Year),nrow = 2) +
  ggtitle("Greater Seattle Area Startups 2010-2019 (companies with no funding removed)") +
  xlab("Number of Founders") +
  ylab("Log 10 of Total Funding to Date in USD") +
  theme(plot.title = element_text(size = 10, face = "bold"), axis.title = element_text(size = 8), axis.text = element_text(size = 7.5))

ggsave("seattle_startups_funding_numoffounders.jpg")

ggplot(median_funding1, aes(x= Categroy.Number.of.Founders, y = Total.Funding.Amount.Currency..in.USD.)) +
    geom_bar(fill = "forestgreen", stat = "identity") + 
    scale_y_continuous(labels = dollar_format(scale = 1/1000000, suffix = "M")) +
    facet_wrap(.~Founded.Year, nrow = 2) +
    coord_flip() +
    ggtitle("Greater Seattle Area Startups 2010-2019") +
    xlab("Number of Founders Category") +
    ylab("Median Total Funding to Date in USD") +
    theme(plot.title = element_text(size = 10, face = "bold"), axis.title = element_text(size = 8), axis.text = element_text(size = 7.5))

ggsave("seattle_startups_medianfunding_numoffounders.jpg")

```

## Step 3 Read all the datasets about the founders and create founder-company join table
```{r}
founder1 <- read.csv("founder_sea_2010to2013.csv", header = T, stringsAsFactors = F)
founder2 <- read.csv("founder_sea_2014to2019.csv", header = T, stringsAsFactors = F)
founders <- rbind(founder1, founder2)

# Error message during bind_rows showed that founder has a column called Numer.of.News.Articles that has comma in numeric number values
founders$Number.of.News.Articles <- as.numeric(gsub(",", "", founders$Number.of.News.Articles))
founders$Number.of.News.Articles <- replace(founders$Number.of.News.Articles, is.na(founders$Number.of.News.Articles),0)
founders <- unique(founders)
```

```{r}
# Join founders table and companies table

names(companies)[2] <- "Primary.Organization.URL"
founder_company <- merge(founders, companies, by ="Primary.Organization.URL", all.x=TRUE)
founder_company

founder_company_experience <- founder_company %>% 
  filter(!is.na(Organization.Name)) %>% 
  group_by(Primary.Organization.URL) %>% 
  mutate(Team.Avg.Num.Founded.Organizations = round(mean(Number.of.Founded.Organizations-1), digits = 1)) %>%
  select(Primary.Organization.URL, Organization.Name, Total.Funding.Amount.Currency..in.USD., Founded.Year, Number.of.Founders, Team.Avg.Num.Founded.Organizations) %>% 
  unique()

# create aggregated median funding table with founder info
median_funding2 <- aggregate(Total.Funding.Amount.Currency..in.USD.~ Founded.Year + Team.Avg.Num.Founded.Organizations, data = founder_company_experience, median)
```

## Step 4 - Data Visualization to examine the relationship between founder experience and funding
```{r}
ggplot(founder_company_experience, aes(x= Team.Avg.Num.Founded.Organizations, y =log10(Total.Funding.Amount.Currency..in.USD. ))) +
  geom_jitter(size = 0.7, color = "forestgreen") +
  geom_smooth(method = "lm", color = "black", lwd = 0.8) +
  facet_wrap(.~as.factor(Founded.Year),nrow = 2) +
  ggtitle("Greater Seattle Area Startups 2010-2019 (companies with no funding removed)") +
  xlab("Average Number of Companies Previsouly Founded per Team") +
  ylab("Log 10 of Total Funding to Date in USD") +
  xlim(0, 3) +
  theme(plot.title = element_text(size = 10, face = "bold"), axis.title = element_text(size = 8), axis.text = element_text(size = 7.5))

ggsave("seattle_startups_funding_serialfounders.jpg")

```

## Step 5 - Data Visualization to examine the relationship between gender diveristy within founding team and funding
```{r}
founder_company_gender <- founder_company %>% 
  filter(!is.na(Organization.Name)) %>% 
  mutate(Gender.Binary = ifelse(Gender == "Male", 1,-1)) %>%
  group_by(Primary.Organization.URL) %>% 
  mutate(Gender.Balance.Index = sum(Gender.Binary)) %>% 
  select(Primary.Organization.URL, Organization.Name, Total.Funding.Amount.Currency..in.USD., Founded.Year, Gender.Balance.Index) %>% 
  unique()

ggplot(founder_company_gender) +
  geom_jitter(aes(x= Gender.Balance.Index, y =log10(Total.Funding.Amount.Currency..in.USD. )), size = 0.7, color = "forestgreen") +
  geom_density(aes(x= Gender.Balance.Index),color = "black") +
  facet_wrap(.~as.factor(Founded.Year),nrow = 2) +
  ggtitle("Greater Seattle Area Startups 2010-2019 (companies with no funding removed)") +
  xlab("Founding Team Gender Balance Index (Number of Male Founders - Number of Female Founders)") +
  ylab("Log 10 of Total Funding to Date in USD") +
  theme(plot.title = element_text(size = 10, face = "bold"), axis.title = element_text(size = 8), axis.text = element_text(size = 7.5))

ggsave("seattle_startups_funding_genderdiversity.jpg")
```



